```{r, include=FALSE}

library (data.table)
library (ggplot2)
library (reshape2)
library (lubridate)
library (stringr)

source ("../R/fetch.R")
source ("../R/features.R")
source ("../R/models.R")

# set global options for all code chunks
opts_chunk$set(fig.width=12, fig.height=8, tidy=T)
```

Kaggle Allstate Purchase Prediction Competition
========================================================

Of course, start out by loading the source data sets.

```{r fetch.data, message=FALSE}
train <- fetch()
```

### Q. What does the data look like?

```{r review.data}
dim (train)
str (train)
summary (train)
```

In the training data, a customer always has a 'purchase' record.  There are no customers in the data who have not purchased a policy.

```{r customers.who.dont.purchase}
customers.no.purchase <- train [ record.type == "purchase", .N, by = customer.id][ N > 1]
dim (customers.no.purchase)
```

### Q. What is the 'shopping.pnt' field?

The 'shopping.pnt' field is incremented each time the customer shops or purchases a policy.  The max value indicates how many contacts the customer had with the insurance company before a policy was purchased.  

For example, here is a customer who had 9 contacts with the insurance company before purchasing a policy.

```{r explore.shopping.pnt}
train [ customer.id == "10000000" ][, c(1:5,19:26), with = F]
```

### Q. How many times does the typical customer interact with Allstate before purchasing a policy?

The unique count of customer interactions across the data set

```{r number.of.customer.interactions}
contacts.by.customer <- train [, .N, by = customer.id ]
sort (unique (contacts.by.customer$N))
```

```{r}
ggplot (contacts.by.customer, aes (N)) + geom_histogram (breaks = 2:14)
```

### Q. Which day of the week do customer contacts occur on?

```{r}
contacts.by.dow <- train [, .N, by = day.of.week]
ggplot (contacts.by.dow, aes(day.of.week, N, fill = day.of.week)) + geom_bar()
```

### Q. When throughout the day do customer contacts occur?

```{r}
ggplot (train, aes(time.of.day.mins)) + geom_density()
ggplot (train, aes(time.of.day.mins)) + geom_density() + facet_wrap( ~ day.of.week, ncol = 4)
```

### Q. A product is composed of a unique selection of the 7 options.  There are 2,304 unique products (3*2*4*3*2*4*4).

```{r}
train[, product := paste0 (option.a, option.b, option.c, option.d, option.e, option.f, option.g)]
```

There are only 1,809 of the 2,304 products in the training set.

```{r}
length (unique (train$product))
```

### Q. We know that cost is a major driver of consumer choices.  If we can attribute a cost for each option choice, we should be able to make some fairly accurate predictions.

```{r}

# transform the options from wide to long
train.melted <- train [, list (customer.id, cost, option.a, 
                               option.b, option.c, option.d, 
                               option.e, option.f, option.g) ]
train.melted <- data.table (melt (train.melted, id.vars = c("customer.id", "cost")))

lim <- quantile (train.melted$cost, c(0.01, 0.99))
boxplot <- ggplot (train.melted, aes (value, cost, fill = value)) + geom_boxplot()
boxplot + scale_y_continuous (limits = lim) + facet_wrap (~variable, ncol = 4, scales="free")

rm (train.melted)
```

### Q. Do customers shop and purchase all on the same day or does this span multiple days?

```{r, echo = F}
shopping.duration (train)
```

```{r view.shopping.duration}

# to find the total/max shopping duration, need to look only at the purchase records
purchases <- train [ record.type == "purchase"]

# build a table
with (purchases, table (shopping.duration))

# build a hist plot
hist <- ggplot (purchases, aes (shopping.duration)) + geom_histogram() 
hist + labs (title = "Duration of Shopping Experience", x = "Duration (Days)")
```

### Q. How frequently do customer attributes change?

*TODO*

### Q. Does group size have a big impact on cost?  Should this be taken into account for option vs cost comparison?

*TODO*

### Q. Which options are most popular for customers?

```{r}

# transform the options from wide to long
train.melted <- train [, list (customer.id, option.a, 
                               option.b, option.c, option.d, 
                               option.e, option.f, option.g) ]
train.melted <- data.table (melt (train.melted, id.vars = c("customer.id")))

# plot it
ggplot (train.melted, aes (value)) + geom_histogram() + facet_wrap (~variable, scales = "free")
```
