Kaggle Allstate Purchase Prediction Competition
========================================================

Of course, start out by loading the source data sets.

<<<<<<< HEAD:src/analysis/exploratory-analysis-nick.Rmd
```{r fetch.data}

source ("../R/fetch.R")
train <- fetch()
=======
```{r fetch.data, message=FALSE}

train <- fetch()

# add some additional features
add.shopping.duration (train)
add.product (train)
```

### Q. What does the data look like?

```{r review.data}
dim (train)
str (train)
summary (train)
```

In the training data, a customer always has a 'purchase' record.  There are no customers in the data who have not purchased a policy.

```{r customers.who.dont.purchase}
customers.no.purchase <- train [ record.type == "purchase", .N, by = customer.id][ N > 1]
dim (customers.no.purchase)
```

### Q. What is the 'shopping.pnt' field?

The 'shopping.pnt' field is incremented each time the customer shops or purchases a policy.  The max value indicates how many contacts the customer had with the insurance company before a policy was purchased.  

For example, here is a customer who had 9 contacts with the insurance company before purchasing a policy.

```{r explore.shopping.pnt}
train [ customer.id == "10000000" ][, 1:5, with = F]
```

### Q. How many times does the typical customer interact with Allstate before purchasing a policy?

The number of times a customer interacts with Allstate before making a purchase.  Of course, the data only contains customers who have made a purchase; not those who have not.

```{r number.of.customer.interactions}
contacts.by.customer <- train [, .N, by = customer.id ]
sort (unique (contacts.by.customer$N))

ggplot (contacts.by.customer, aes (N)) + geom_histogram (breaks = 2:14)
```

### Q. Which day of the week do customer contacts occur on?

```{r}
ggplot (train, aes(day.of.week, fill = day.of.week)) + geom_histogram()
```

### Q. When throughout the day do customer contacts occur?

```{r}
ggplot (train, aes(time.of.day.hours)) + geom_density()
ggplot (train, aes(time.of.day.hours)) + geom_density() + facet_wrap( ~ day.of.week, ncol = 4)
```

### Q. A product is composed of a unique selection of the 7 options.  There are 2,304 unique products (3*2*4*3*2*4*4).

There are only 1,809 of the 2,304 available 'products' in the training set.

```{r}
length (unique (train$product))
```

### Q. We know that cost is a major driver of consumer choices.  If we can attribute a cost for each option choice, we should be able to make some fairly accurate predictions.

```{r, message=F}

# transform the options from wide to long
train.melted <- train [, list (customer.id, cost, option.a, 
                               option.b, option.c, option.d, 
                               option.e, option.f, option.g) ]
train.melted <- data.table (melt (train.melted, id.vars = c("customer.id", "cost")))

lim <- quantile (train.melted$cost, c(0.01, 0.99))
p <- ggplot (train.melted, aes (value, cost, fill = value)) + geom_boxplot()
p + scale_y_continuous (limits = lim) + facet_wrap (~variable, ncol = 4, scales = "free")
```

### Q. Do customers shop and purchase all on the same day or does this span multiple days?

```{r view.shopping.duration}

# to find the total/max shopping duration, need to look only at the purchase records
purchases <- train [ record.type == "purchase"]

# build a table
with (purchases, table (shopping.duration))
>>>>>>> 0f42bb5eff03770fd6f7fc151000cf9ec56494a9:src/analysis/explore-nick.Rmd

```

<<<<<<< HEAD:src/analysis/exploratory-analysis-nick.Rmd
=======
### Q. Does group size have a big impact on cost?  Should this be taken into account for option vs cost comparison?

There are many more samples in the data of smaller groups.  The group size in the training data is skewed to the right.

```{r}
ggplot (train, aes (as.factor (group.size))) + geom_histogram()
```

The cost does not change much based on the group size.  I would expect the cost to increase dramatically as the group size increases.  Perhaps this is not the case with insurance or the 'cost' is really a cost per individual.

There also appears to be more outliers the smaller the group is.  This is probably due to there being very few samples of groups of size 3, 4+.

```{r}
ggplot (train, aes (as.factor (group.size), cost)) + geom_boxplot()
```

### Q. Which options are most popular for customers?

```{r}

# transform the options from wide to long
train.melted <- train [, list (customer.id, option.a, 
                               option.b, option.c, option.d, 
                               option.e, option.f, option.g) ]
train.melted <- data.table (melt (train.melted, id.vars = c("customer.id")))

# plot it
p <- ggplot (train.melted, aes (value)) + geom_histogram (aes (fill = value)) 
p + facet_wrap (~variable, scales = "free") + theme (legend.position = "none")
```

### Q. How frequently do customer attributes change?

*TODO*
>>>>>>> 0f42bb5eff03770fd6f7fc151000cf9ec56494a9:src/analysis/explore-nick.Rmd
