
```{r, include = FALSE}

library (recommenderlab)
library (data.table)
library (caret)
library (stringr)

source ("../R/fetch.R")
source ("../R/features.R")
source ("../R/score.R")
source ("../R/utils.R")
source ("../R/cache.R")
source ("../R/customer-model.R")

```

Recommender
========================================================

Can user-based collaborative filtering help predict a customer's purchase?

 - Item -> Plan Options
 - User -> Customer
 - User Attributes -> State, Location, Car Age, etc, etc
 - Ratings -> The number of times a user shopped for a particular option. A user's actual purchase might be weighted to indicate a much higher rating.

The available recommenders.

```{r}
recommenderRegistry$get_entries (dataType = "realRatingMatrix")
```


```{r}

#
# load the raw data
#

# if a user has shopping for an option N times this counts as a +N rating
raw <- fetch ()
quotes <- extract.quotes (raw, sum)
extract.purchases (raw, quotes)

#
# focus only on option a for now
#

# the columns specific to the option being analyzed; option a in this case
keep.columns <- c("customer.id", na.omit (str_match (names (quotes), "option\\.a.*")))
quotes <- quotes [, keep.columns, with = FALSE]

#
# create the ratings matrix
#

# extract the ratings matrix from the data
ratings <- as.matrix (subset (quotes, select = 2:(ncol (quotes)-1)))
rownames (ratings) <- quotes$customer.id

# the recommende expects NA to indicate "no rating".  all 0s need to be replaced NA
#ratings [ratings == 0] <- NA
ratings <- as (ratings, "realRatingMatrix")

# normalize the ratings - this should help account for bias caused by some shoppers who received more quotes
#ratings <- normalize (ratings)

# TODO - are 'goodRating' and 'given' parameters correct?

# train 
eval <- evaluationScheme (ratings, method = "split", train = 0.8, goodRating = 1, given = 1)
recommender <- Recommender (getData (eval, "train"), method = "POPULAR")

# predict 
p <- predict (recommender, getData (eval, "known"), type = "ratings")

# TODO - choose the option with the maximum predicted rating

# TODO - score

```

If we can use purchases data from the training set, we might weight that more heavily and it would look something like the following.  I don't think there will be any training with this approach though.

```{r}

# extract pseudo-ratings from the purchase data
purchases <- raw [ record.type == "purchase" ]

# transform the options into dummy variables; option.a becomes (option.a0, option.a1, option.a2)
dummies <- dummyVars (~ option.a + option.b + option.c + option.d + option.e + option.f + option.g, purchases)
purchases <- data.table (customer.id = purchases$customer.id, predict (dummies, purchases))

# summarize the options which were shopped for using summarize.func
purchases <- purchases [, lapply(.SD, summarize.func, ...), by = customer.id ]
setkey (purchases, "customer.id")

# if a user has purchased an option this counts as a +5 rating
purchases <- purchases [, keep.columns, with = FALSE ]

option.columns <- keep.columns [2:length(keep.columns)]
options.only <- purchases [, option.columns, with = FALSE]

purchase.weight <- 5
options.only <- options.only * purchase.weight

# TODO - add back the customer id and convert back to data table

```



