```{r, include=FALSE}

library (data.table)
library (ggplot2)
library (reshape2)
library (lubridate)
library (stringr)

source ("../R/fetch.R")
source ("../R/features.R")
source ("../R/models.R")
source ("../R/score.R")
source ("../R/utils.R")

# set global options for all code chunks
opts_chunk$set(fig.width=12, fig.height=8, tidy=T)
```

Modeling of Customer Shopping History
========================================================

The customer demographic data and the customer shopping history are completely different types of data.  We are going to model each of these independently.  The final prediction will be an ensemble between the two.  The initial challenge for the shopping history data is in flattening a customer's shopping history into one single record that can be used to train and predict the customer's final purchase plan.  

### Preprocessing

The input shopping data looks like the following.

| customer.id | shopping.pnt | record.type | day.of.week | time.of.day.hours | option.a | option.b | ... |
| ----------- | ------------ | ----------- | ----------- | ----------------- | -------- | -------- | ----|                                                 
| 123456789   | 1            | shopping    | Mon         | 2.25              | 0        | 1        |     |
| 123456789   | 2            | shopping    | Mon         | 3.25              | 0        | 1        |     |
| 123456789   | 3            | shopping    | Mon         | 3.50              | 1        | 2        |     |
| 123456789   | 4            | shopping    | Mon         | 4.00              | 1        | 2        |     |
| 123456789   | 5            | shopping    | Mon         | 5.00              | 0        | 2        |     |
| 123456789   | 6            | purchase    | Mon         | 6.26              | 0        | 2        |     |
| 000000001   | 1            | shopping    | Tue         | 2.25              | 0        | 1        |     |
| 000000001   | 2            | shopping    | Tue         | 3.25              | 0        | 1        |     |
| 000000001   | 3            | purchase    | Tue         | 3.50              | 1        | 2        |     |

The input data could be transformed as follows for training and prediction.

| customer.id | option.a.0 | option.a.1 | option.a.2 | option.b.0 | option.b.1 | option.a | option.b | ... | 
| ----------- | ---------- | ---------- | ---------- | ---------- | ---------- | -------- | -------- | --- |
| 123456789   | 3          | 2          | 0          | 2          | 2          | 0        | 2        |     |
| 000000001   | 2          | 0          | 0          | 0          | 2          | 1        | 2        |     |

The definition for each of these fields is as follows.

| Field         | Definition 
| ------------- | ---------- 
| *customer.id* | The customer's unique ID.  Each customer will have one record in training data.       
| *option.a.0*  | The total number of times the customer shopped for choice 0 of option A in the shopping history.
| *option.a.1*  | ...
| *option.a.2*  | ...
| *option.b.0*  | The total number of times the customer shopped for choice 0 of option A in the shopping history.
| *option.b.1*  | ...
| *option.a*    | The outcome for option a that is being predicted.  The value for this field comes from the purchase record.
| *option.b*    | ...

Does this work??

```{r}
train <- fetch()
shopping <- flatten.shopping.history (train)
str (shopping)
```


### Modeling

#### TODO: This takes care of the option elements of the shopping history, but how should the additional date-related 'day.of.week' and 'time.of.day.hours' fields be leveraged in the analysis.

#### TODO: Perhaps a difference value for each of the 'option.x.n' fields could be used.  Perhaps ones that applies a greater weight to more recent shopping history.





